---
title: "Mini-Projet"
author: "Vincent Dufresne, Benjamin Gosselin"
output: 
  html_notebook: 
    toc: yes
    toc_float: yes
    collapsed: false
    smooth_scroll: false
    toc_depth: 4
---


# Installing dependencies
```{r}
#install.packages("tseries")
#install.packages("tsbox")
#install.packages("psych")
library(tseries)
library(tsbox)
library(psych)
```


# Data Extraction
```{r}
xic <- get.hist.quote(instrument = "XIC.TO", start = "2013-04-15", end = "2022-03-11", quote = "AdjClose")
xus <- get.hist.quote(instrument = "XUS.TO", start = "2013-04-15", end = "2022-03-11", quote = "AdjClose")
```
# Manipulating Data Frames

We will first need to generate log returns and we also need to split data between estimation set and "out-of-sample"
set which we will later use to analyze the performance the model.

## Generating Log Returns
```{r}
#We generate log returns for all observations of XIC and XUS
xic$logreturn <- log(xic$Adjusted)

xus$logreturn <- log(xus$Adjusted)
```

## Creating data frames

### XIC
```{r}
xic_df <- ts_df(xic[,1])
colnames(xic_df) <- c("Date", "Adjusted")
xic_df$logreturn <- log(xic_df$Adjusted)
colnames(xic_df) <- c("Date", "Adjusted", "logreturn")
head(xic_df)

```

### XUS
```{r}
xus_df <- ts_df(xus[,1])
colnames(xus_df) <- c("Date", "Adjusted")
xus_df$logreturn <- log(xus_df$Adjusted)
colnames(xus_df) <- c("Date", "Adjusted", "logreturn")
head(xus_df)
```



## Splitting data

```{r}
#Estimation set size
est_size = 2000

#Test set size
test_size <- length(xic[,1]) - est_size
```


### XIC
```{r}
#Estimation set
xic_est <- xic[1:est_size,]
xic_est_df <- xic_df[1:est_size,]

#Test set
xic_test <- xic[-(1:est_size),]
xic_test_df <- xic_df[-(1:est_size),]

#Checking lengths are ok
length(xic_est[,1]) + length(xic_test[,1]) == length(xic[,1])

```



### XUS
```{r}
#Estimation set
xus_est <- xus[1:est_size,]
xus_est_df <- xus_df[1:est_size,]

#Test set
xus_test <- xus[-(1:est_size),]
xus_test_df <- xus_df[-(1:est_size),]

#Checking lengths are ok
length(xus_est[,1]) + length(xus_test[,1]) == length(xus[,1])

```

# Analyzing Data

## Summary of series

### XUS

#### Full set
```{r}
describe(xic[, 1:2])
plot(xic)
par(mfrow=c(1,2))
plot(x=xic_df$Date, y=xic_df$Adjusted, type='l', xlab="Date", ylab="Adjusted", main="Ajdusted as a function of Date")
plot(x=xic_df$Date, y=xic_df$logreturn, type='l', xlab="Date", ylab="logreturn", main="logreturn as a function of Date")

```


### XIC

#### Full set
```{r}
describe(xus[,1:2])
plot(xus)
```


#### Estimation set
```{r}
describe(xus_est[,1:2])
plot(xus_est)
```

#### Test set
```{r}
describe(xus_test[,1:2])
plot(xus_test)
```




```{r}

```

