---
title: "Mini-Projet"
author: "Vincent Dufresne, Benjamin Gosselin"
output: 
  html_notebook: 
    toc: yes
    toc_float: yes
    collapsed: false
    smooth_scroll: false
    toc_depth: 4
---


# Installing dependencies
```{r}
#install.packages("tseries")
#install.packages("tsbox")
#install.packages("psych")
library(xts)
library(tseries)
library(tsbox)
library(psych)
library(copula)
```


# Data Extraction
```{r}
xic <- get.hist.quote(instrument = "XIC.TO", start = "2013-04-15", end = "2022-03-11", quote = "AdjClose")
xus <- get.hist.quote(instrument = "XUS.TO", start = "2013-04-15", end = "2022-03-11", quote = "AdjClose")
```
# Manipulating Data Frames

We will first need to generate log returns and we also need to split data between estimation set and "out-of-sample"
set which we will later use to analyze the performance the model.

## Generating Log Returns
```{r}
#We generate log returns for all observations of XIC and XUS
xic <- ts_df(xic)
xus <- ts_df(xus)

colnames(xic) <- c("Date", "Adjusted")
colnames(xus) <- c("Date", "Adjusted")

xic$logprice <- log(xic$Adjusted)
xus$logprice <- log(xus$Adjusted)

xic_log_ret <- c()

for (i in 1:(length(xic$Adjusted)-1)){
  log_ret <- as.numeric(xic[i+1,2]) - as.numeric(xic[i,2])
  xic_log_ret <- c(xic_log_ret, log_ret)
}

xic_log_ret <- c(0, xic_log_ret)
xic$logret <- xic_log_ret

xic

xus_log_ret <- c()

for (i in 1:(length(xus$Adjusted)-1)){
  log_ret <- as.numeric(xus[i+1,2]) - as.numeric(xus[i,2])
  xus_log_ret <- c(xus_log_ret, log_ret)
}

xus_log_ret <- c(0, xus_log_ret)

xus$logret <- xus_log_ret

xus

```



## Splitting data

```{r}
#Estimation set size
est_size = 2000

#Test set size
test_size <- length(xic[,4]) - est_size
```


### XIC
```{r}
#Estimation set
xic_est <- xic[1:est_size,]
xic_est_df <- xic[1:est_size,]

#Test set
xic_test <- xic[-(1:est_size),]
xic_test_df <- xic[-(1:est_size),]

#Checking lengths are ok
length(xic_est[,1]) + length(xic_test[,1]) == length(xic[,1])

```



### XUS
```{r}
#Estimation set
xus_est <- xus[1:est_size,]
xus_est_df <- xus[1:est_size,]

#Test set
xus_test <- xus[-(1:est_size),]
xus_test_df <- xus[-(1:est_size),]

#Checking lengths are ok
length(xus_est[,1]) + length(xus_test[,1]) == length(xus[,1])

```

# Analyzing Data

## Summary of series

### XIC

#### Full set
```{r}
desc <- describe(xic[, c(2,4)])
desc$excess_kur <- desc$kurtosis - 3
desc
plot(xic)

plot(x=xic$Date, y=xic$logret, type='l', xlab="Date", ylab="logreturn", main="logreturn as a function of Date")

```

#### Estimation set
```{r}
desc <- describe(xic_est[, c(2,4)])
desc$excess_kur <- desc$kurtosis - 3
desc

plot(xic_est)

plot(x=xic_est_df$Date, y=xic_est_df$logret, type='l', xlab="Date", ylab="logreturn", main="logreturn as a function of Date")
```



#### Test set
```{r}
desc <- describe(xic_test[, 1:2])
desc$excess_kur <- desc$kurtosis - 3
desc

plot(xic_test)

plot(x=xic_test_df$Date, y=xic_test_df$logret, type='l', xlab="Date", ylab="logreturn", main="logreturn as a function of Date")

```



### XUS

#### Full set
```{r}
desc <- describe(xus[,1:2])
desc$excess_kur <- desc$kurtosis - 3
desc

plot(xus)

plot(x=xus$Date, y=xus$logret, type='l', xlab="Date", ylab="logreturn", main="logreturn as a function of Date")
```


#### Estimation set
```{r}
desc <- describe(xus_est[, 1:2])
desc$excess_kur <- desc$kurtosis - 3
desc

plot(xus_est)

plot(x=xus_est$Date, y=xus_est$logret, type='l', xlab="Date", ylab="logreturn", main="logreturn as a function of Date")
```

#### Test set
```{r}
desc <- describe(xus_test[, 1:2])
desc$excess_kur <- desc$kurtosis - 3
desc
plot(xus_test)

plot(x=xus_test$Date, y=xus_test$logret, type='l', xlab="Date", ylab="logreturn", main="logreturn as a function of Date")
```

# Models

## Gaussian

$$r_{i,t} = \mu_i + \sigma_{i,t}\epsilon_{i,t},\space where \space \epsilon_{i,t} \sim N(0,1)$$

### Model Estimation

We know R has a function to model garch models for data of ts type (time series) so we will use it. To get more info on how to use the function one can run the next chunk of code
```{r}
#install.packages("rugarch")
library(rugarch)
library(xts)
?rugarch
```

#### Model Specifications
```{r}
ug_gauss <- ugarchspec()
```



Once we are familiarized, we can fit our model

#### XIC

```{r}
class(xic_est)

lr_xic <- as.xts(xic_est$logret, order.by = xic_est$Date)
ugfit_xic <- ugarchfit(spec = ug_gauss, data = lr_xic)

ugfit_xic

plot(ugfit_xic@fit$fitted.values)

ug_xic_var <- ugfit_xic@fit$var
ug_xic_res2 <- (ugfit_xic@fit$residuals)^2

plot(ug_xic_res2)
lines(ug_xic_var, col='green')
```




#### XUS

```{r}
lr_xus <- as.xts(xus_est$logret, order.by = xus_est$Date)
ugfit_xus <- ugarchfit(spec = ug_gauss, data = lr_xus)

ugfit_xus

ugfit_xus@fit$coef
ug_xus_var <- ugfit_xus@fit$var
ug_xus_res2 <- (ugfit_xus@fit$residuals)^2

plot(ug_xus_res2)
lines(ug_xus_var, col='green')
```



### Copulas

```{r}
plot(xic$logret, xus$logret, xlab="xic", ylab="xus")
cor(xic$logret, xus$logret, method="kendall")
```

```{r}
p_xic_res <- pobs(ugfit_xic@fit$residuals)
p_xus_res <- pobs(ugfit_xus@fit$residuals)

p_res <- data.frame(p_xic_res)

p_res$p_xus_res <- p_xus_res

p_res

cor(p_res, method="kendall")
```



#### Gaussian

```{r}
g_cop <- normalCopula(param = 0.1, dim = 2)
g_cop

gauss_cop_ml <- fitCopula(g_cop, data = p_res, method="ml")
gauss_cop_mpl <- fitCopula(g_cop, data = p_res, method="mpl")
gauss_cop_ml
gauss_cop_mpl
```


#### Student-t

```{r}
t_cop = tCopula(param = 0.1, dim = 2, df = 2, df.fixed = TRUE)
t_cop

student_cop_ml <- fitCopula(t_cop, data = p_res, method = "ml")
student_cop_mpl <- fitCopula(t_cop, data = p_res, method = "mpl")

student_cop_ml
student_cop_mpl
```


#### Frank

```{r}
f_cop <- frankCopula(param = 0.1, dim = 2)

frank_cop_ml <- fitCopula(f_cop, data = p_res, method = "ml")
frank_cop_mpl <- fitCopula(f_cop, data = p_res, method = "mpl")

frank_cop_ml
frank_cop_mpl
```


#### Clayton

```{r}
clay_cop <- claytonCopula(param = 2, dim = 2)

clayton_cop_ml <- fitCopula(clay_cop, data = p_res, method = "ml")
clayton_cop_mpl <- fitCopula(clay_cop, data = p_res, method = "mpl")

clayton_cop_ml
clayton_cop_mpl
```


## Student-t

$$r_{i,t} = \mu_i + \sigma_{i,t}\epsilon_{i,t},\space where \space \epsilon_{i,t} \sim t_{\nu i}$$


```{r}
ugspec_t <- ugarchspec(distribution.model = "std")
ugspec_t
```

#### XIC
```{r}
ugfit_t_xic <- ugarchfit(spec=ugspec_t, data=lr_xic)

ugfit_t_xic
```


#### XUS
```{r}
ugfit_t_xus <- ugarchfit(spec = ugspec_t, data = lr_xus)

ugfit_t_xus
```

### Copulas


```{r}
p_xic_t_res <- pobs(ugfit_t_xic@fit$residuals)
p_xus_t_res <- pobs(ugfit_t_xus@fit$residuals)

p_res_t <- data.frame(p_xic_t_res)

p_res_t$p_xus_res <- p_xus_t_res

plot(p_res_t)

cor(p_res_t, method="kendall")
```


#### Gaussian

```{r}
gauss_cop_t = normalCopula(param = 0.1, dim = 2)

gaussian_cop_t_ml <- fitCopula(gauss_cop_t, data = p_res_t, method = "ml")
gaussian_cop_t_mpl <- fitCopula(gauss_cop_t, data = p_res_t, method = "mpl")

gaussian_cop_t_ml
gaussian_cop_t_mpl
```

#### Student-t

```{r}
t_cop_t <- tCopula(param = 0.1, dim = 2, df = 2, df.fixed = TRUE)

student_cop_ml_t <- fitCopula(t_cop_t, data = p_res_t, method = "ml")
student_cop_ml_t
```


#### Frank

```{r}
fr_cop_t <- frankCopula(param = 0.1, dim = 2)

frank_cop_ml_t <- fitCopula(fr_cop_t, data = p_res_t, method = "ml")
frank_cop_ml_t
```


#### Clayton

```{r}
cl_cop_t <- claytonCopula(param = 0.1, dim = 2)

clayton_cop_ml_t <- fitCopula(cl_cop_t, data = p_res_t, method = "ml")
clayton_cop_ml_t

llike <- logLik(clayton_cop_ml_t)
AIC <- 2*1 - 2*as.numeric(llike)
AIC

BIC <- log(length(p_res_t[,1])) - 2*as.numeric(llike)
BIC
```



# VaR Forecasting

## Estimation portfolio

### Graphical VaR
```{r}
library(ggplot2)

w_xic <- 0.5
w_xus <- 1 - w_xic

ptf_est <- data.frame(lr_xic)

ptf_est$xus <- lr_xus

colnames(ptf_est) <- c("xic", "xus")

ptf_est$ptf_ret <- w_xic * ptf_est$xic + w_xus * ptf_est$xus

ptf_est

par(mfrow=c(1,2))

qplot(ptf_est$ptf_ret , geom = 'histogram') + geom_histogram(fill = 'lightblue' , bins = 30) +
    geom_histogram(aes(ptf_est$ptf_ret[ptf_est$ptf_ret < quantile(ptf_est$ptf_ret , 0.05)]) , fill = 'red' , bins = 30) +
    labs(x = 'Daily logReturns')

library(ggplot2)
qplot(ptf_est$ptf_ret , geom = 'histogram') + geom_histogram(fill = 'lightblue' , bins = 30) +
    geom_histogram(aes(ptf_est$ptf_ret[ptf_est$ptf_ret < quantile(ptf_est$ptf_ret , 0.01)]) , fill = 'red' , bins = 30) +
    labs(x = 'Daily logReturns')

```

